Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-11-19 doc/Paper/main.tex
Started implementing comments from Jo.

      \item {\bf Survey depth}: we estimate the $10\sigma$ survey depth in different directions as follows. For each pixel, we compute the average $i$-band flux error of all objects classified as stars that fall into it. The corresponding survey depth is then given by this mean flux error multiplied by 10 and translated into a magnitude. Not all pixels contain enough stars to estimate a reliable mean flux error. We use a nearest-neighbor interpolation to assign a depth value to those pixels that contain fewer than 4 stars. We verified that we obtain similar depth estimates using other methods, such as that outlined in \cite{2018PASJ...70S..25M}, in which 10$\sigma$ depth is defined as the mean {\tt cmodel} magnitude of all sources with $i$-band signal-to-noise ratio between 9 and 11. We emphasize that our method uses only stars, discarding all sources classified as galaxies. The use of galaxies to estimate any systematics map is dangerous in the context of template deprojection, since the stochastic noise in the map associated with the discrete sampling of these sources correlates directly with the galaxy distribution, and can give rise to spurious correlations that bias the estimated power spectrum after deprojection (see Section \ref{ssec:methods.cell}). We provide further details in Appendix \ref{app:depth}.

@@

    Note that an underlying assumption of this step is that the coadded mean of different exposures fully captures the connection between fluctuations in observing conditions and the artificial fluctuations they cause on the number counts. In general we could also consider other cumulants of the per-exposure distribution of observing condition values. As shown below, we do not observe an important contamination from these systematics in our results, and we therefore leave this more thorough study for future work.

@@

  \subsection{Angular power spectra}\label{ssec:methods.cell}
    We compute angular power spectra using a flat-sky pseudo-$C_\ell$ (PCL) algorithm \citep{2002ApJ...567....2H} as implemented in {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}}. The reader is referred to the code's paper \cite{2019MNRAS.484.4127A} for a detailed description of the estimator, but we provide a brief summary here for completeness.

@@

    As noted above, the power spectra are computed in rings of $\ell=|{\bf l}|$, which we will call bandpowers here. We use 17 contiguous bandpowers with edges $\ell=$(100, 200, 300, 400, 600, 800, 1000, 1400, 1800, 2200, 3000, 3800, 4600, 6200, 7800, 9400, 12600, 158800). Due to this bandpower averaging, the estimated power spectra cannot, strictly speaking, be compared with theoretical predictions estimated at token multipoles (e.g. the midpoint of each band). The effect of this averaging can however be taken into account exactly as a linear operation of the form $\bar{C}_b=\sum_\ell F^b_{\ell} C_\ell$, where $C_\ell$ is the theoretical prediction evaluated at all integer multipoles $\ell$, $\bar{C}_b$ is the prediction for the $b$-th bandpower and the bandpower windows $F^b_{\ell}$ incorporate the effects of mode-coupling, averaging into bandpowers and the inversion of the binned mode-coupling matrix.

@@

  \subsubsection{Covariance matrices}\label{sssec:methods.theory.covar}
    We use an analytical procedure to estimate the uncertainties of our measured power spectra, inspired by the methods used by \cite{Krause:2017}. As shown in \cite{2009MNRAS.395.2065T,Takada:2013}, the covariance matrix for large-scale structure data can be decomposed into a disconnected trispectrum part, essentially equivalent to the covariance of a Gaussian random field with the same power spectrum as the data, a connected part, caused by the non-Gaussian nature of the density field, and a super-sample covariance term (labelled SSC here). The SSC contribution accounts for the coherent shift in the amplitude of density fluctuations within the surveyed volume caused by long wavelength modes larger than the survey.

@@

  \subsection{Parameter constraints}\label{ssec:methods.constr}
    In order to derive constraints on HOD, cosmological and systematics parameters, we assume the joint likelihood of all auto- and cross-power spectra to be Gaussian 
    \begin{align}
      \mathscr{L}(D \vert \theta) = \frac{1}{[(2\pi)^{d}\det{\mathsf{C}}]^{\sfrac{1}{2}}} e^{-\frac{1}{2}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})^{\mathrm{T}}\mathsf{C}^{-1}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})}.
      \label{eq:likelihood}
    \end{align}
    The quantity $\mathsf{C}$ denotes the joint covariance matrix, which we estimate analytically as described in Sec.~\ref{sssec:methods.theory.covar} and which we keep constant during parameter estimation \citep{2019OJAp....2E...3K}.
    We sample the likelihood in a Monte Carlo Markov Chain (MCMC) using the publicly available code \texttt{CosmoHammer} \cite{Akeret:2013}, which is based on \texttt{emcee} \cite{Foreman-Mackey2013}\footnote{We note that we have compared the MCMC results obtained using \texttt{CosmoHammer} to those obtained using a Metropolis-Hastings algorithm as implemented in \texttt{april} (https://github.com/slosar/april), finding consistent results for our test case.}. In our fiducial analysis we sample the parameter set $\boldsymbol{\theta} = \{\mu_{\mathrm{min}}, \allowbreak \, \mu_{\mathrm{min}, p}, \allowbreak \, \mu_{0}, \allowbreak \, \mu_{0, p}, \allowbreak \, \mu_{1}, \allowbreak \, \mu_{1, p}, \allowbreak \, \Delta z_{i}, \allowbreak \, z_{w, i}\}$, $i = 0, \dots ,3$, where the first six parameters describe the HOD of galaxies as outlined in Sec.~\ref{sssec:methods.theory.hod}. The remaining parameters account for photometric redshift uncertainties as described in Sec.~\ref{sssec:methods.theory.photoz_syst}. We perform additional analyses in which we separately allow for variations in the amplitude $A_{\mu}$ of the magnification bias kernel $W_{\mu}$ and the cosmological parameters $\Omega_{c}$ and $\sigma_{8}$, where $\Omega_{c}$ is the fractional cold matter density today and $\sigma_{8}$ denotes the r.m.s. of linear matter fluctuations in spheres of comoving radius 8 $h^{-1}$ Mpc. For all sampled parameters, we assume flat, uniform priors. The sampled parameters are shown alongside their priors in Tab.~\ref{tab:params}. The remaining HOD parameters are set to $\alpha = 1.0$ and $\sigma_{\log M} = 0.4$. Unless stated otherwise, we further fix all cosmological parameters to the best-fit values derived by the Planck Collaboration in 2018 using temperature, polarization and CMB lensing data, i.e. $\Omega_{b}=0.0493$, $\Omega_{c}=0.264$, $h=0.6736$, $n_{s}=0.9649$ and $\sigma_{8}=0.8111$ (see the fourth column in Tab.~2 in \cite{Planck:2018}).

@@

    \subsubsection{Interpretation of HOD results}\label{sssec:results.hod-constraints.constraints-interpretation}
      As described in Sec.~\ref{sssec:results.hod-constraints.fiducial}, our best-fit HOD constraints predict both the minimal mass to host a central galaxy $M_{\mathrm{min}}(z)$ and the mass scale for satellites $M_{1}(z)$ to be redshift independent. In order to understand this trend, it would be ideal to look at the rest-frame properties of the galaxies in our sample and how these change with redshift. However, computing k-corrections and absolute magnitudes for broad-band photometric data can be challenging and we therefore choose an alternative approach. We cross-match galaxies passing our selection criteria within the HSC COSMOS field to galaxies also included in the COSMOS 30-band photometric catalog of \cite{2016ApJS..224...24L}. The COSMOS 30-band catalog contains, amongst others, absolute magnitudes in Subaru \texttt{B}-band, $M_{B}$, and Subaru \texttt{r+}-band, $M_{R}$. Using these quantities, we construct color-magnitude diagrams for all cross-matched galaxies. Fig.~\ref{fig:color-mag} shows the $M_{B}-M_{R}$ color as a function of $M_{R}$ magnitude for 10 equally spaced redshift bins in $z \in [0.15, 1.5]$\footnote{The galaxies are split into bins according to COSMOS 30-band photometric redshifts. We note that we have repeated the analysis splitting the sample according to HSC \texttt{Ephor\_AB} photometric redshifts, finding consistent results.}. In this diagram, red galaxies populate the high $M_{B}-M_{R}$ color, high $M_{R}$ magnitude plane, while blue galaxies tend to have lower $M_{B}-M_{R}$ and $M_{R}$. As can be seen from the figure, red galaxies increasingly drop-out of our sample at high redshift. We can quantify this effect by computing the fraction of red galaxies in our sample as a function of redshift: we empirically determine the separation between the blue and red clouds (denoted by the solid lines in the figure) and compute the fraction of galaxies in the red cloud as $f_{R} = \sfrac{N_{R}}{(N_{R}+N_{B})}$, where $N_{R}$ denotes the number of galaxies above the separation line, while $N_{B}$ is the number of galaxies below the line. Consistent with the figure, we find a decreasing $f_{R}$ as a function of redshift.

@@

\section{Discussion and Conclusions}\label{sec:discussion}
  In this work, we have presented a comprehensive analysis of photometric galaxy clustering in HSC-SSP DR1 data using a magnitude-limited sample with $i<24.5$. We have split this sample into four tomographic redshift bins and computed all possible auto- and cross-power spectra, accounting for observational systematics in number density fluctuations through map-level deprojection. We have then fitted the data using a halo model coupled with a halo occupation distribution, marginalizing over a flexible model for photometric redshift uncertainties. Our results are well fitted by our theoretical model including scales up to $k_{\mathrm{max}} = 1$ Mpc$^{-1}$. We find monotonically decreasing average halo masses, which can be interpreted in terms of the drop-out of red galaxies at high redshifts for a magnitude-limited sample, consistent with previous analyses. In order to test the stability of our results, we performed a suite of alternative analyses, finding our results to be robust to photometric redshift uncertainties. We extended our analysis to include the effects of lensing magnification, which results in a $\sim 3 \sigma$ detection of magnification from our data alone. In addition to this, we also allowed for variations in the $\Lambda$CDM cosmological parameters, finding that the data are able to constrain $\Omega_{c}$, while $\sigma_{8}$ remains unconstrained. Finally, we used our data to derive a simple fitting function for the linear, large-scale galaxy bias of magnitude-limited samples as a function of redshift and limiting $i$-band magnitude (Eq. \ref{eq:bias-fit-func-result}).

>>>>

      \item {\bf Survey depth}: we estimate the $10\sigma$ survey depth as a function of angular position on the sky as follows. For each pixel, we compute the average $i$-band flux error of all objects classified as stars that fall into it. The corresponding survey depth is then given by this mean flux error multiplied by 10 and translated into a magnitude. Not all pixels contain enough stars to estimate a reliable mean flux error. We use a nearest-neighbor interpolation to assign a depth value to those pixels that contain fewer than 4 stars. We verified that we obtain similar depth estimates using other methods, such as that outlined in \cite{2018PASJ...70S..25M}, in which 10$\sigma$ depth is defined as the mean {\tt cmodel} magnitude of all sources with $i$-band signal-to-noise ratio between 9 and 11. We emphasize that our method uses only stars, discarding all sources classified as galaxies. The use of galaxies to estimate any systematics map is dangerous in the context of template deprojection, since the stochastic noise in the map associated with the discrete sampling of these sources correlates directly with the galaxy distribution, and can give rise to spurious correlations that bias the estimated power spectrum after deprojection (see Section \ref{ssec:methods.cell}). We provide further details in Appendix \ref{app:depth}.

@@

    Note that an underlying assumption of this step is that the coadded mean of different exposures fully captures the connection between fluctuations in observing conditions and the artificial fluctuations they cause on the number counts. In general we could also consider other cumulants of the per-exposure distribution of observing condition values. As shown in Sec.~\ref{sssec:results.spectra.syst}, we do not observe an important contamination from these systematics in our results, and we therefore leave this more thorough study for future work.

@@

  \subsection{Angular power spectra}\label{ssec:methods.cell}
    We compute angular power spectra using a flat-sky pseudo-$C_\ell$ (PCL) algorithm \citep{2002ApJ...567....2H} as implemented in {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}.}. The reader is referred to the code's paper \cite{2019MNRAS.484.4127A} for a detailed description of the estimator, but we provide a brief summary here for completeness.

@@

    As noted above, the power spectra are computed in rings of $\ell=|{\bf l}|$, which we will call bandpowers here. We use 17 contiguous bandpowers with edges $\ell=$(100, 200, 300, 400, 600, 800, 1000, 1400, 1800, 2200, 3000, 3800, 4600, 6200, 7800, 9400, 12600, 15800). Due to this bandpower averaging, the estimated power spectra cannot, strictly speaking, be compared with theoretical predictions estimated at token multipoles (e.g. the midpoint of each band). The effect of this averaging can however be taken into account exactly as a linear operation of the form $\bar{C}_b=\sum_\ell F^b_{\ell} C_\ell$, where $C_\ell$ is the theoretical prediction evaluated at all integer multipoles $\ell$, $\bar{C}_b$ is the prediction for the $b$-th bandpower and the bandpower windows $F^b_{\ell}$ incorporate the effects of mode-coupling, averaging into bandpowers and the inversion of the binned mode-coupling matrix.

@@

  \subsubsection{Covariance matrices}\label{sssec:methods.theory.covar}
    We use an analytical procedure to estimate the uncertainties of our measured power spectra, inspired by the methods used by \cite{Krause:2017}. As shown in \cite{2009MNRAS.395.2065T,Takada:2013}, the covariance matrix for large-scale structure data can be decomposed into a disconnected trispectrum part, essentially equivalent to the covariance of a Gaussian random field with the same power spectrum as the data, a connected part, caused by the non-Gaussian nature of the density field, and a super-sample covariance term (labelled SSC here)\footnote{A comparison of the relative contributions of these three terms to the total covariance can be found in Fig.~\ref{fig:covariance-contributions}.}. The SSC contribution accounts for the coherent shift in the amplitude of density fluctuations within the surveyed volume caused by long wavelength modes larger than the survey.

@@

  \subsection{Parameter constraints}\label{ssec:methods.constr}
    In order to derive constraints on HOD, cosmological and systematics parameters, we assume the joint likelihood of all auto- and cross-power spectra to be Gaussian 
    \begin{align}
      \mathscr{L}(D \vert \theta) = \frac{1}{[(2\pi)^{d}\det{\mathsf{C}}]^{\sfrac{1}{2}}} e^{-\frac{1}{2}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})^{\mathrm{T}}\mathsf{C}^{-1}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})}.
      \label{eq:likelihood}
    \end{align}
    The quantity $\mathsf{C}$ denotes the joint covariance matrix, which we estimate analytically as described in Sec.~\ref{sssec:methods.theory.covar} and which we keep constant during parameter estimation \citep{2019OJAp....2E...3K}.
    We sample the likelihood in a Monte Carlo Markov Chain (MCMC) using the publicly available code \texttt{CosmoHammer} \cite{Akeret:2013}, which is based on \texttt{emcee} \cite{Foreman-Mackey2013}\footnote{We note that we have compared the MCMC results obtained using \texttt{CosmoHammer} to those obtained using a Metropolis-Hastings algorithm as implemented in \texttt{april} (https://github.com/slosar/april), finding consistent results for our test case.}. In our fiducial analysis we sample the parameter set $\boldsymbol{\theta} = \{\mu_{\mathrm{min}}, \allowbreak \, \mu_{\mathrm{min}, p}, \allowbreak \, \mu_{0}, \allowbreak \, \mu_{0, p}, \allowbreak \, \mu_{1}, \allowbreak \, \mu_{1, p}, \allowbreak \, \Delta z_{i}, \allowbreak \, z_{w, i}\}$, $i = 0, \dots ,3$, where the first six parameters describe the HOD of galaxies as outlined in Sec.~\ref{sssec:methods.theory.hod}. The remaining parameters account for photometric redshift uncertainties as described in Sec.~\ref{sssec:methods.theory.photoz_syst}. We perform additional analyses in which we separately allow for variations in the amplitude $A_{\mu}$ of the magnification bias kernel $W_{\mu}$ and the cosmological parameters $\Omega_{c}$ and $\sigma_{8}$, where $\Omega_{c}$ is the fractional cold matter density today and $\sigma_{8}$ denotes the r.m.s. of linear matter fluctuations in spheres of comoving radius 8 $h^{-1}$ Mpc. For all sampled parameters, we assume flat, uniform priors. The sampled parameters are shown alongside their priors in Tab.~\ref{tab:params}\footnote{The values of these parameters derived from our analysis will be discussed in Sec.~\ref{ssec:results.hod-constraints}.}. The remaining HOD parameters are set to $\alpha = 1.0$ and $\sigma_{\log M} = 0.4$, consistent with the simulation results of Ref.~\cite{2005ApJ...633..791Z}. Unless stated otherwise, we further fix all cosmological parameters to the best-fit values derived by the Planck Collaboration in 2018 using temperature, polarization and CMB lensing data, i.e. $\Omega_{b}=0.0493$, $\Omega_{c}=0.264$, $h=0.6736$, $n_{s}=0.9649$ and $\sigma_{8}=0.8111$ (see the fourth column in Tab.~2 in \cite{Planck:2018}).

@@

    \subsubsection{Interpretation of HOD results}\label{sssec:results.hod-constraints.constraints-interpretation}
      As described in Sec.~\ref{sssec:results.hod-constraints.fiducial}, our best-fit HOD constraints predict both the minimal mass to host a central galaxy $M_{\mathrm{min}}(z)$ and the mass scale for satellites $M_{1}(z)$ to be redshift independent. As noted above, this is somewhat counter-intuitive, as we would expect at least $M_{\mathrm{min}}(z)$ to increase with redshift. In order to understand this trend, it would be ideal to look at the rest-frame properties of the galaxies in our sample and how these change with redshift. However, computing k-corrections (see e.g. \cite{Hogg:2002, Blanton:2007}) and absolute magnitudes for broad-band photometric data can be challenging and we therefore choose an alternative approach. We cross-match galaxies passing our selection criteria within the HSC COSMOS field to galaxies also included in the COSMOS 30-band photometric catalog of \cite{2016ApJS..224...24L}. The COSMOS 30-band catalog contains, amongst others, absolute magnitudes in Subaru \texttt{B}-band, $M_{B}$, and Subaru \texttt{r+}-band, $M_{R}$. Using these quantities, we construct color-magnitude diagrams for all cross-matched galaxies. Fig.~\ref{fig:color-mag} shows the $M_{B}-M_{R}$ color as a function of $M_{R}$ magnitude for 10 equally spaced redshift bins in $z \in [0.15, 1.5]$\footnote{The galaxies are split into bins according to COSMOS 30-band photometric redshifts. We note that we have repeated the analysis splitting the sample according to HSC \texttt{Ephor\_AB} photometric redshifts, finding consistent results.}. In this diagram, red galaxies populate the high $M_{B}-M_{R}$ color, high $M_{R}$ magnitude plane, while blue galaxies tend to have lower $M_{B}-M_{R}$ and $M_{R}$. As can be seen from the figure, red galaxies increasingly drop-out of our sample at high redshift. We can quantify this effect by computing the fraction of red galaxies in our sample as a function of redshift: we empirically determine the separation between the blue and red clouds (denoted by the solid lines in the figure) and compute the fraction of galaxies in the red cloud as $f_{R} = \sfrac{N_{R}}{(N_{R}+N_{B})}$, where $N_{R}$ denotes the number of galaxies above the separation line, while $N_{B}$ is the number of galaxies below the line. Consistent with the figure, we find a decreasing $f_{R}$ as a function of redshift.

@@

\section{Discussion and Conclusions}\label{sec:discussion}
  In this work, we have presented a comprehensive analysis of photometric galaxy clustering in HSC-SSP DR1 data using a magnitude-limited sample with $i<24.5$. We have split this sample into four tomographic redshift bins and computed all possible auto- and cross-power spectra, accounting for observational systematics in number density fluctuations through map-level deprojection. We have then fitted the data using a halo model coupled with a halo occupation distribution, marginalizing over a flexible model for photometric redshift uncertainties. Our results are well fitted by our theoretical model including scales up to $k_{\mathrm{max}} = 1$ Mpc$^{-1}$. We find monotonically decreasing average halo masses as a function of redshift, which can be interpreted in terms of the drop-out of red galaxies at high redshifts for a magnitude-limited sample, consistent with previous analyses. In order to test the stability of our results, we performed a suite of alternative analyses, finding our results to be robust to photometric redshift uncertainties. We extended our analysis to include the effects of lensing magnification, which results in a $\sim 3 \sigma$ detection of magnification from our data alone. In addition to this, we also allowed for variations in the $\Lambda$CDM cosmological parameters, finding that the data are able to constrain $\Omega_{c}$, while $\sigma_{8}$ remains unconstrained. Finally, we used our data to derive a simple fitting function for the linear, large-scale galaxy bias of magnitude-limited samples as a function of redshift and limiting $i$-band magnitude (Eq. \ref{eq:bias-fit-func-result}).
